The lattice of acyclic pipe dreams

### 2022-10-17 acyclicFacets/draft/latticePipeDreams.tex
start algorithms

% pipe dreams
\input{pipeDreamsTikz}
% cross
%\newcommand{\cross}[1][black]{\raisebox{.1cm}{\pipeDreamMonoColor{#1}{black}{c}}}
\newcommand{\cross}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{cross}}}
\newcommand{\crossBiColor}[2]{\raisebox{.1cm}{\pipeDreamBiColor{#1}{#2}{black}{c/l/r}}}
\newcommand{\NScross}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{NScross}}}
\newcommand{\WEcross}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{WEcross}}}
% elbow
%\newcommand{\elbow}[1][black]{\raisebox{.1cm}{\pipeDreamMonoColor{#1}{black}{e}}}
\newcommand{\elbow}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{elbow}}}
\newcommand{\elbowBiColor}[2]{\raisebox{.1cm}{\pipeDreamBiColor{#1}{#2}{black}{e/l/r}}}
%\newcommand{\SEelbow}[1][black]{\raisebox{.1cm}{\pipeDreamBiColor{white}{#1}{black}{e/l/r}}}
\newcommand{\SEelbow}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{SEelbow}}}
%\newcommand{\WNelbow}[1][black]{\raisebox{.1cm}{\pipeDreamBiColor{#1}{white}{black}{e/l/r}}}
\newcommand{\WNelbow}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{WNelbow}}}
\newcommand{\waving}{\mathsf{wav}}%{\raisebox{-.2cm}{\begin{tikzpicture}[scale=1]\pipeDreamBiColor{.2}{(0,0)}{black}{white}{{t/r/l},{t/l/l},{t/l/r}}\end{tikzpicture}}}
\newcommand{\transpose}{\mathsf{transp}}
\newcommand{\pipeDreams}{\Pi} % set of pipe dreams
\newcommand{\HP}{{\bk\pipeDreams}} % Hopf Algebra of pipe dreams
\newcommand{\reversingPipeDreams}{\Omega} % reversing pipe dreams
\newcommand{\contact}{^\#} % contact graph
\newcommand{\se}{\textsc{se}} % southeast
\newcommand{\nw}{\textsc{nw}} % northwest
\newcommand{\insertion}[2]{\mathsf{ins}(#1,#2)} % insertion of permutation #1 in pipe dreams with permutation #2
\newcommand{\duality}{^\star} % duality pipe dreams - triangulations
\newcommand{\acyclicPipeDreams}{\Sigma} % acyclic pipe dreams
\newcommand{\linearExtensions}{\mathcal{L}} % linear extensions
\newcommand{\noninversions}[2]{\mathsf{ninv}(#1,#2)} % non-inversions
\newcommand{\acyclicOrientations}{\Omega} % set of pipe dreams
\newcommand{\recoils}[2]{\mathsf{rec}(#1,#2)} % recoil scheme of a pipe dream
\newcommand{\canopy}[1]{\mathsf{can}(#1)} % canopy of a pipe dream
\newcommand{\greedyPipeDream}{P^{\overleftarrow{gr}}} % greedy facet
\newcommand{\antiGreedyPipeDream}{P^{\overrightarrow{gr}}} % antigreedy facet

@@

A \defn{pipe dream}~$P$ is a filling of a triangular shape with crosses~\cross{} and elbows~\elbow{} so that all pipes entering on the left side exit on the top side.
We only consider \defn{reduced} pipe dreams, where two pipes have at most one intersection.
We label the pipes with~$1, 2, \dots, n$ in the order of their entry points from top to bottom.
We denote by~$\omega_P \in \fS_n$ the order of the exit points of the pipes of~$P$ from left to right.
In other words, the pipe entering at row~$i$ exits at column~$\omega^{-1}_P(i)$.
For a fixed permutation~$\omega \in \fS_n$, we denote by~$\pipeDreams(\omega)$ the set of reduced pipe dreams~$P$ such that~$\omega_P = \omega$.
% We let~$\pipeDreams_n \eqdef \bigsqcup_{\omega \in \fS_n} \pipeDreams(\omega)$ and~$\pipeDreams \eqdef \bigsqcup_{n \in \N} \pipeDreams_n$.

\begin{figure}[ht]
	\centerline{
		%\input{pipeDreams}
		\includegraphics[scale=.9]{pipeDreams}
	}
	\caption{Two pipe dreams of~$\pipeDreams(1365724)$ connected by a flip (exchanging an elbow with the crossing on the two red pipes~$3$ and~$4$).}
	\label{fig:pipeDreams}
\end{figure}

Two pipe dreams are related by a \defn{flip} if they define the same permutation of the pipes and only differ by the position of a cross and an elbow.
An elbow~$e$ in a pipe dream~$P$ is \defn{flippable} if the two pipes passing through elbow~$e$ have a crossing~$c$, and the flip exchanges the elbow~$e$ with the cross~$c$.
See \cref{fig:pipeDreams}.
The flip is \defn{increasing} if the elbow~$e$ is weakly southwest of the crossing~$c$.
For example, the flip of \cref{fig:pipeDreams} is increasing from left to right.
%
The \defn{increasing flip graph} is the graph of increasing flips on~$\pipeDreams(\omega)$.
It is clearly a directed acyclic graph, and it has a unique source and a unique sink \cite{PilaudPocchiola}, called the \defn{greedy} and \defn{antigreedy} pipe dreams, and denoted~$\greedyPipeDream$ and~$\antiGreedyPipeDream$.
\vincent{In \cite{PilaudStump-ELlabeling}, these are called positive and negative greedy facets. Also the notations are different...}
The \defn{increasing flip poset} is the reflexive and transitive closure of the increasing flip graph on~$\pipeDreams(\omega)$.

The \defn{contact graph} of a pipe dream~$P$ is the directed graph~$P\contact$ with one node for each pipe of~$P$ and one arc for each elbow of~$P$ connecting the northwest pipe to the southeast pipe of the elbow\footnote{We have reversed the usual orientation conventions of \cite{PilaudSantos-brickPolytope, PilaudPocchiola, Pilaud-BrickAlgebra} to suit better our purposes, in particular in \cref{subsec:insertionAlgorithm}.}.
We see equivalently the contact graph~$P\contact$ as a graph on the pipes of~$P$ or on the integers~$[n]$.
We say that a pipe dream~$P$ is \defn{acyclic} if its contact graph~$P\contact$ is (no oriented cycle).
We then denote by~$\contactLess{P}$ the transitive closure of the contact graph of~$P$.
For~$\omega \in \fS_n$, we denote by~$\acyclicPipeDreams(\omega)$ the set of acyclic pipe dreams of~$\pipeDreams(\omega)$.
% We let~$\acyclicPipeDreams_n \eqdef \bigsqcup_{\omega \in \fS_n} \acyclicPipeDreams(\omega)$ and~$\acyclicPipeDreams \eqdef \bigsqcup_{n \in \N} \acyclicPipeDreams_n$.
%\vincent{Add contact graphs to \cref{fig:pipeDreams}.}

\begin{example}
\label{exm:Tamari1}
We say that a pipe dream is \defn{reversing} if it fixes the first and last pipes and reverses the order of the remaining pipes.
In this case, it is natural to label the pipes from~$0$ to~$n+1$, so that the permutation of the pipes is~$\rho_n \eqdef [0, n, n-1, \dots, 2, 1, n+1]$.
% We denote by~$\reversingPipeDreams_n$ the reversing pipe dreams of~$\pipeDreams_n$.
As observed in~\cite{Woo, PilaudPocchiola, Pilaud-these, Stump}, the family of reversing pipe dreams belong to the Catalan families.
\cref{fig:bijection} illustrates explicit bijections between reversing pipe dreams of~$\pipeDreams(\rho_n)$, binary trees with $n$ internal nodes, and the triangulations of a convex $(n+2)$-gon.
More precisely, the map which sends an elbow in row~$i$ and column~$j$ of the triangular shape to the diagonal~$[i,n+1-j]$ of the~$(n+2)$-gon provides the following correspondence: \\[.3cm]
\centerline{
\begin{tabular}{r@{$\quad\longleftrightarrow\quad$}l}
pipe dream~$P \in \reversingPipeDreams_n$ & triangulation~$P\duality$ of the $(n+2)$-gon, \\
pipe $i$ of~$P$ & triangle $i\duality$ of~$P\duality$ (with central vertex~$i$), \\
elbow between pipes~$i$ and~$j$ of~$P$ & common side of triangles~$i\duality$ and~$j\duality$ of~$P\duality$, \\
cross between pipes~$i$ and~$j$ of~$P$ & common bisector between triangles~$i\duality$ and~$j\duality$ of~$P\duality$, \\
contact graph of~$P$ & dual binary tree of~$P\duality$, \\
elbow flips in~$P$ & diagonal flips in~$P\duality$. \\
%increasing flip poset & Tamari lattice.
\end{tabular}} \\[.3cm]
%
\begin{figure}
	\centerline{\includegraphics[scale=1.3]{dualiteTriangulation}}
	\caption{The bijection between reversing pipe dreams (left), binary trees (middle) and triangulations (right).}
	\label{fig:bijection}
\end{figure}
%
As a consequence, this map sends the increasing flip poset on reversing pipe dreams to the \defn{Tamari lattice} on binary trees.
This lattice is defined as the transitive closure of right rotations on binary trees, and is obtained as the quotient of the weak order by the \defn{sylvester congruence}.
\vincent{Not sure where we want that...}
An important point here is that all reversing pipe dreams are acyclic, since there contact graphs are (oriented) binary trees.
\end{example}

@@

\begin{lemma}
\label{lem:characterizationPipeDreams}
A collection~$P$ of~$n$ pipes pairwise disjoint except at crossing and elbows and such that for each~$j \in [n]$, the pipe~$j$ enters at row~$j$, exits at column~$\omega^{-1}(j)$, and has~$\noninversions{\omega}{j}$ southeast elbows is a pipe dream of~$\pipeDreams(\omega)$.
\end{lemma}

@@

\begin{proof}
Let~$x$ (resp.~$y$) be the location of an elbow of pipe~$i$ (resp.~$j$) such that~$x$ is weakly northwest of~$y$.
We proceed by induction on the grid distance from~$x$ to~$y$.
If they coincide, then pipes~$i$ and~$j$ share an elbow, so that there is an edge from~$i$ to~$j$ in~$P\contact$.
Otherwise, let~$k$ be the pipe of~$P$ with a southeast elbow at~$x$ ($k$ is either the pipe~$i$ itself, or there is an edge from~$i$ to~$k$ in~$P\contact$) and~$\ell$ be the pipe of~$P$ with a northwest elbow at~$y$ ($\ell$ is either the pipe~$j$ itself, or there is an edge from~$\ell$ to~$j$ in~$P\contact$).
Let~$R$ be the axis-parallel rectangle with elbows~$x$ and~$y$.
Since pipes~$k$ and~$\ell$ cross at most once, at least one of them has an additional elbow along the sides of~$R$.
Assume for instance that~$k$ has an elbow at~$x'$.
Then~$x'$ is still weakly northwest of~$y$ and $x',y$ are strictly closer than~$x,y$.
By induction, there is a directed path from~$k$ to~$\ell$ in~$P\contact$, and thus a path directed form~$i$ to~$j$.
\end{proof}

@@

\begin{proof}
Let~$e$ denote the contact of pipes~$i$ and~$k$ in~$P$.
Decompose the triangular shape into three regions: the region~$A$ of all points located southwest of~$e$, the region~$B$ of all points located northwest or southeast of~$e$, and the region~$C$ of all points located northeast of~$e$.
Since~$i < j$ and~$\omega^{-1}(j) > \omega^{-1}(k)$, the pipe~$j$ starts in region~$A$ and ends in region~$C$.
Hence, the pipe~$j$ has an elbow~$f$ in region~$B$.
We thus obtain that~$i \contactLess{P} j \contactMore{P} k$ if this $f$ is southeast of~$e$, and that~$i \contactMore{P} j \contactLess{P} k$ if $f$ is northwest of~$e$.
\end{proof}

%\begin{lemma}
%\label{lem:consequenceRectangle2}
%Let~$P$ be a pipe dream and let~$i,j,k$ be three pipes of~$P$ such that~$\min(i,k) < j$ and~$\min(\omega^{-1}(i), \omega^{-1}(k)) < \omega^{-1}(j)$.
%If~$i \to k$ in~$P\contact$, then either~$i \contactLess{P} j \contactMore{P} k$ or~$i \contactMore{P} j \contactLess{P} k$.
%\vincent{not sure we really use this extended version}
%\end{lemma}
%
%\begin{proof}
%Let~$e$ denote the contact of pipes~$i$ and~$k$ in~$P$.
%Decompose the triangular shape into three regions: the region~$A$ of all points located southwest of~$e$, the region~$B$ of all points located northwest or southeast of~$e$, and the region~$C$ of all points located northeast of~$e$.
%Since~$\min(i,k) < j$ and~$\min(\omega^{-1}(i), \omega^{-1}(k)) < \omega^{-1}(j)$, the pipe~$j$ starts in region~$A$ and ends in region~$C$.
%Hence, the pipe~$j$ has an elbow~$f$ in region~$B$.
%We thus obtain that~$i \contactLess{P} j \contactMore{P} k$ if this $f$ is southeast of~$e$, and that~$i \contactMore{P} j \contactLess{P} k$ if $f$ is northwest of~$e$.
%\end{proof}

@@

\begin{proof}
For any~$i < j$ with~$\omega^{-1}(i) < \omega^{-1}(j)$, the pipes~$i$ and~$j$ do not cross.
Hence, the pipe~$j$ passes southeast of every elbow of the pipe~$i$.
This shows that~$i \contactLess{P} j$ by \cref{lem:rectangle}, thus ${\pi^{-1}(i) < \pi^{-1}(j)}$.
In other words, any non-inversion of~$\omega$ is a non-inversion of~$\pi$, so that~${\pi < \omega}$.
\end{proof}

@@

\begin{proof}
We just need to show that the poset~$\contactLess{P}$ (given by the transitive closure of~$P\contact$) satisfies the conditions of \cref{prop:WOIP}.
Consider~$i < j < k$ such that~$i \contactLess{P} k$.
If~$\omega^{-1}(i) < \omega^{-1}(j)$, then~$i \contactLess{P} j$ by \cref{lem:consequenceRectangle1}.
Similarly, if~$\omega^{-1}(j) < \omega^{-1}(k)$, then~$j \contactLess{P} k$ by \cref{lem:consequenceRectangle1}.
We can therefore assume that~$\omega^{-1}(i) > \omega^{-1}(j) > \omega^{-1}(k)$.
Decompose the triangular shape into three regions: the region~$A$ of all points located northeast of the last elbow of the pipe~$j$ of~$P$, the region~$B$ of all points located northwest or southeast of an elbow of the pipe~$j$ of~$P$, and the region~$C$ of all points located southwest of the first elbow of the pipe~$j$ of~$P$.
Since~$i \contactLess{P} k$, there is a path~$\pi$ form the exiting point of the pipe~$i$ of~$P$ to the entering point of the pipe~$j$ of~$P$ which travels along the pipes of~$P$, possibly jumping from the northwest pipe to the southeast pipe of an elbow it encounters.
Since~$i < j$ and~$\omega^{-1}(j) > \omega^{-1}(k)$, the path~$\pi$ starts in region~$A$ and ends in region~$C$, so that it necessarily passes from region~$A$ to region~$C$.
Since the southwest corner of~$A$ is located northeast of the northeast corner of~$C$, this forces an elbow~$e$ of~$\pi$ to lie in region~$B$.
\cref{lem:rectangle} then ensures that either~$i \contactLess{P} j$ (if~$e$ is above the pipe~$j$), or~$j \contactLess{P} k$ (if~$e$ is below the pipe~$j$).
The proof is similar if~$i \contactMore{P} k$.
\end{proof}

@@

\begin{lemma}
\label{lem:flippable}
Consider two acyclic pipe dreams~$P,P' \in \acyclicPipeDreams(\omega)$ connected by the flip of an elbow between their pipes~$i$ and~$j$.
Then any directed path in~$P\contact$ or~$P'{}\contact$ between~$i$ and~$j$ is an arc.
\end{lemma}

@@

\vincent{This section is in progress}
In this section, we describe three algorithms to decide whether two permutations are equivalent for the pipe dream congruence~$\equiv_\omega$.

@@

We now define a surjective map from the permutations of~$[e, \omega]$ to the pipe dreams of~$\acyclicPipeDreams(\omega)$.
This map, adapted from~\cite{Pilaud-BrickAlgebra}, is obtained by an insertion process similar to the insertion in binary search trees.
%To properly describe this insertion algorithm, we introduce additional notations.
We now describe this insertion algorithm, which is illustrated in \cref{fig:insertionAlgorithm}.

Consider a given permutation~$\pi \eqdef \pi(1) \dots \pi(n)$ (written in one-line notation) in~$[e, \omega]$.
We start from the empty triangular shape and insert the pipes~$\pi(1), \dots, \pi(n)$ one by one in the order of the permutation~$\pi$ as northwest as possible.
In other words, at step~$t$, we consider the set~$X_t$ of free northwest elbows in the rectangle~$[\pi(t)] \times [\omega^{-1}(\pi(t))]$.
The pipe inserted at step~$t$ then starts in row~$\pi(t)$, ends in column~$\omega^{-1}(\pi(t))$, covers each elbow of~$X_t$ with a southeast elbow, and adds just one new free northwest elbow in between any two consecutive elbows of~$X_t$.
We denote the resulting pipe dream by~$\insertion{\pi}{\omega}$.
See \cref{fig:insertionAlgorithm}.

\begin{figure}[b]
	\centerline{
		\includegraphics[scale=.9]{insertion}
	}
	\caption{The insertion~$\insertion{513264}{561324}$.}
	\label{fig:insertionAlgorithm}
\end{figure}

>>>>

% pipe dreams
\input{pipeDreamsTikz}
% cross
%\newcommand{\cross}[1][black]{\raisebox{.1cm}{\pipeDreamMonoColor{#1}{black}{c}}}
\newcommand{\cross}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{cross}}}
\newcommand{\crossBiColor}[2]{\raisebox{.1cm}{\pipeDreamBiColor{#1}{#2}{black}{c/l/r}}}
\newcommand{\NScross}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{NScross}}}
\newcommand{\WEcross}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{WEcross}}}
% elbows
%\newcommand{\elbow}[1][black]{\raisebox{.1cm}{\pipeDreamMonoColor{#1}{black}{e}}}
\newcommand{\elbow}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{elbow}}}
\newcommand{\elbowBiColor}[2]{\raisebox{.1cm}{\pipeDreamBiColor{#1}{#2}{black}{e/l/r}}}
%\newcommand{\SEelbow}[1][black]{\raisebox{.1cm}{\pipeDreamBiColor{white}{#1}{black}{e/l/r}}}
\newcommand{\SEelbow}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{SEelbow}}}
%\newcommand{\WNelbow}[1][black]{\raisebox{.1cm}{\pipeDreamBiColor{#1}{white}{black}{e/l/r}}}
\newcommand{\WNelbow}[1][black]{\raisebox{-.15cm}{\includegraphics[scale=.9]{WNelbow}}}
\newcommand{\waving}{\mathsf{wav}}%{\raisebox{-.2cm}{\begin{tikzpicture}[scale=1]\pipeDreamBiColor{.2}{(0,0)}{black}{white}{{t/r/l},{t/l/l},{t/l/r}}\end{tikzpicture}}}
\newcommand{\transpose}{\mathsf{transp}}
\newcommand{\pipeDreams}{\Pi} % set of pipe dreams
\newcommand{\HP}{{\bk\pipeDreams}} % Hopf Algebra of pipe dreams
\newcommand{\reversingPipeDreams}{\Omega} % reversing pipe dreams
\newcommand{\contact}{^\#} % contact graph
\newcommand{\se}{\textsc{se}} % southeast
\newcommand{\nw}{\textsc{nw}} % northwest
\newcommand{\insertion}[2]{\mathsf{ins}(#1,#2)} % insertion of permutation #1 in pipe dreams with permutation #2
\newcommand{\duality}{^\star} % duality pipe dreams - triangulations
\newcommand{\acyclicPipeDreams}{\Sigma} % acyclic pipe dreams
\newcommand{\linearExtensions}{\mathcal{L}} % linear extensions
\newcommand{\noninversions}[2]{\mathsf{ninv}(#1,#2)} % non-inversions
\newcommand{\acyclicOrientations}{\Omega} % set of pipe dreams
\newcommand{\recoils}[2]{\mathsf{rec}(#1,#2)} % recoil scheme of a pipe dream
\newcommand{\canopy}[1]{\mathsf{can}(#1)} % canopy of a pipe dream
\newcommand{\greedyPipeDream}{P^{\overleftarrow{gr}}} % greedy facet
\newcommand{\antiGreedyPipeDream}{P^{\overrightarrow{gr}}} % antigreedy facet

@@

A \defn{pipe dream}~$P$ is a filling of a triangular shape with crossings~\cross{} and contacts~\elbow{} so that all pipes entering on the left side exit on the top side.
We only consider \defn{reduced} pipe dreams, where two pipes have at most one crossing.
We label the pipes with~$1, 2, \dots, n$ in the order of their entry points from top to bottom.
We denote by~$\omega_P \in \fS_n$ the order of the exit points of the pipes of~$P$ from left to right.
In other words, the pipe entering at row~$i$ exits at column~$\omega^{-1}_P(i)$.
For a fixed permutation~$\omega \in \fS_n$, we denote by~$\pipeDreams(\omega)$ the set of reduced pipe dreams~$P$ such that~$\omega_P = \omega$.
% We let~$\pipeDreams_n \eqdef \bigsqcup_{\omega \in \fS_n} \pipeDreams(\omega)$ and~$\pipeDreams \eqdef \bigsqcup_{n \in \N} \pipeDreams_n$.

\begin{figure}[ht]
	\centerline{
		%\input{pipeDreams}
		\includegraphics[scale=.9]{pipeDreams}
	}
	\caption{Two pipe dreams of~$\pipeDreams(1365724)$ connected by a flip (exchanging a contact with the crossing on the two red pipes~$3$ and~$4$).}
	\label{fig:pipeDreams}
\end{figure}

%Two pipe dreams are related by a \defn{flip} if they only differ by the position of a crossing and a contact (in particular, they define the same permutation).
A contact~$c$ is \defn{flippable} if the two pipes passing through contact~$c$ have a crossing~$x$.
The \defn{flip} exchanges the contact~$c$ with the crossing~$x$.
%See \cref{fig:pipeDreams}.
The flip is \defn{increasing} if the contact~$c$ is weakly southwest of the crossing~$x$.
For example, \cref{fig:pipeDreams} illustrates an increasing from the left pipe dream to the right pipe dream.
%
The \defn{increasing flip graph} is the graph of increasing flips on~$\pipeDreams(\omega)$.
It is clearly a directed acyclic graph, and it has a unique source and a unique sink \cite{PilaudPocchiola}, called the \defn{greedy} and \defn{antigreedy} pipe dreams, and denoted~$\greedyPipeDream$ and~$\antiGreedyPipeDream$.
\vincent{In \cite{PilaudStump-ELlabeling}, these are called positive and negative greedy facets. Also the notations are different...}
The \defn{increasing flip poset} is the reflexive and transitive closure of the increasing flip graph on~$\pipeDreams(\omega)$.

The \defn{contact graph} of a pipe dream~$P$ is the directed graph~$P\contact$ with one node for each pipe of~$P$ and one arc for each contact of~$P$ connecting the northwest pipe to the southeast pipe of the contact\footnote{We have reversed the usual orientation conventions of \cite{PilaudSantos-brickPolytope, PilaudPocchiola, Pilaud-BrickAlgebra} to suit better our purposes, in particular in \cref{subsec:insertionAlgorithm}.}.
We see equivalently the contact graph~$P\contact$ as a graph on the pipes of~$P$ or on the integers~$[n]$.
We say that a pipe dream~$P$ is \defn{acyclic} if its contact graph~$P\contact$ is acyclic (no oriented cycle).
We then denote by~$\contactLess{P}$ the transitive closure of the contact graph of~$P$.
For~$\omega \in \fS_n$, we denote by~$\acyclicPipeDreams(\omega)$ the set of acyclic pipe dreams of~$\pipeDreams(\omega)$.
% We let~$\acyclicPipeDreams_n \eqdef \bigsqcup_{\omega \in \fS_n} \acyclicPipeDreams(\omega)$ and~$\acyclicPipeDreams \eqdef \bigsqcup_{n \in \N} \acyclicPipeDreams_n$.
%\vincent{Add contact graphs to \cref{fig:pipeDreams}.}

\begin{example}
\label{exm:Tamari1}
We say that a pipe dream is \defn{reversing} if it fixes the first and last pipes and reverses the order of the remaining pipes.
In this case, it is natural to label the pipes from~$0$ to~$n+1$, so that the permutation of the pipes is~$\rho_n \eqdef [0, n, n-1, \dots, 2, 1, n+1]$.
% We denote by~$\reversingPipeDreams_n$ the reversing pipe dreams of~$\pipeDreams_n$.
As observed in~\cite{Woo, PilaudPocchiola, Pilaud-these, Stump}, the family of reversing pipe dreams belong to the Catalan families.
\cref{fig:bijection} illustrates explicit bijections between reversing pipe dreams of~$\pipeDreams(\rho_n)$, binary trees with $n$ internal nodes, and the triangulations of a convex $(n+2)$-gon.
More precisely, the map which sends a contact in row~$i$ and column~$j$ of the triangular shape to the diagonal~$[i,n+1-j]$ of the~$(n+2)$-gon provides the following correspondence: \\[.3cm]
\centerline{
\begin{tabular}{r@{$\quad\longleftrightarrow\quad$}l}
pipe dream~$P \in \reversingPipeDreams_n$ & triangulation~$P\duality$ of the $(n+2)$-gon, \\
pipe $i$ of~$P$ & triangle $i\duality$ of~$P\duality$ (with central vertex~$i$), \\
contact between pipes~$i$ and~$j$ of~$P$ & common side of triangles~$i\duality$ and~$j\duality$ of~$P\duality$, \\
crossing between pipes~$i$ and~$j$ of~$P$ & common bisector of triangles~$i\duality$ and~$j\duality$ of~$P\duality$, \\
contact graph of~$P$ & dual binary tree of~$P\duality$, \\
contact flips in~$P$ & diagonal flips in~$P\duality$. \\
%increasing flip poset & Tamari lattice.
\end{tabular}} \\[.3cm]
%
\begin{figure}
	\centerline{\includegraphics[scale=1.3]{dualiteTriangulation}}
	\caption{The bijection between reversing pipe dreams (left), binary trees (middle) and triangulations (right).}
	\label{fig:bijection}
\end{figure}
%
As a consequence, this map sends the increasing flip poset on reversing pipe dreams to the \defn{Tamari lattice} on binary trees.
This lattice is defined as the transitive closure of right rotations on binary trees, and is obtained as the quotient of the weak order by the \defn{sylvester congruence}.
\vincent{Not sure where we want that...}
An important point here is that all reversing pipe dreams are acyclic, since there contact graphs are (oriented) binary trees.
\end{example}

@@

\begin{lemma}
\label{lem:characterizationPipeDreams}
A collection~$P$ of~$n$ pipes pairwise disjoint except at crossing and contacts and such that for each~$j \in [n]$, the pipe~$j$ enters at row~$j$, exits at column~$\omega^{-1}(j)$, and has~$\noninversions{\omega}{j}$ southeast contacts is a pipe dream of~$\pipeDreams(\omega)$.
\end{lemma}

@@

\begin{proof}
Let~$x$ (resp.~$y$) be the location of an elbow of pipe~$i$ (resp.~$j$) such that~$x$ is weakly northwest of~$y$.
We proceed by induction on the grid distance from~$x$ to~$y$.
If they coincide, then pipes~$i$ and~$j$ share a contact, so that there is an edge from~$i$ to~$j$ in~$P\contact$.
Otherwise, let~$k$ be the pipe of~$P$ with a southeast elbow at~$x$ ($k$ is either the pipe~$i$ itself, or there is an edge from~$i$ to~$k$ in~$P\contact$) and~$\ell$ be the pipe of~$P$ with a northwest elbow at~$y$ ($\ell$ is either the pipe~$j$ itself, or there is an edge from~$\ell$ to~$j$ in~$P\contact$).
Let~$R$ be the axis-parallel rectangle with corners~$x$ and~$y$.
Since pipes~$k$ and~$\ell$ cross at most once, at least one of them has an additional elbow along the sides of~$R$.
Assume for instance that~$k$ has an elbow at~$x'$.
Then~$x'$ is still weakly northwest of~$y$ and $x',y$ are strictly closer than~$x,y$.
By induction, there is a directed path from~$k$ to~$\ell$ in~$P\contact$, and thus a path directed form~$i$ to~$j$.
\end{proof}

@@

\begin{proof}
Let~$c$ denote the contact of pipes~$i$ and~$k$ in~$P$.
Decompose the triangular shape into three regions: the region~$A$ of all points located southwest of~$c$, the region~$B$ of all points located northwest or southeast of~$c$, and the region~$C$ of all points located northeast of~$c$.
Since~$i < j$ and~$\omega^{-1}(j) > \omega^{-1}(k)$, the pipe~$j$ starts in region~$A$ and ends in region~$C$.
Hence, the pipe~$j$ has an elbow~$e$ in region~$B$.
We thus obtain that~$i \contactLess{P} j \contactMore{P} k$ if this $e$ is southeast of~$c$, and that~$i \contactMore{P} j \contactLess{P} k$ if $e$ is northwest of~$c$.
\end{proof}

%\begin{lemma}
%\label{lem:consequenceRectangle2}
%Let~$P$ be a pipe dream and let~$i,j,k$ be three pipes of~$P$ such that~$\min(i,k) < j$ and~$\min(\omega^{-1}(i), \omega^{-1}(k)) < \omega^{-1}(j)$.
%If~$i \to k$ in~$P\contact$, then either~$i \contactLess{P} j \contactMore{P} k$ or~$i \contactMore{P} j \contactLess{P} k$.
%\vincent{not sure we really use this extended version}
%\end{lemma}
%
%\begin{proof}
%Let~$e$ denote the contact of pipes~$i$ and~$k$ in~$P$.
%Decompose the triangular shape into three regions: the region~$A$ of all points located southwest of~$e$, the region~$B$ of all points located northwest or southeast of~$e$, and the region~$C$ of all points located northeast of~$e$.
%Since~$\min(i,k) < j$ and~$\min(\omega^{-1}(i), \omega^{-1}(k)) < \omega^{-1}(j)$, the pipe~$j$ starts in region~$A$ and ends in region~$C$.
%Hence, the pipe~$j$ has a contact~$f$ in region~$B$.
%We thus obtain that~$i \contactLess{P} j \contactMore{P} k$ if this $f$ is southeast of~$e$, and that~$i \contactMore{P} j \contactLess{P} k$ if $f$ is northwest of~$e$.
%\end{proof}

@@

\begin{proof}
For any~$i < j$ with~$\omega^{-1}(i) < \omega^{-1}(j)$, we have~$i \contactLess{P} j$ by \cref{lem:consequenceRectangle1}, thus ${\pi^{-1}(i) < \pi^{-1}(j)}$ since~$\pi \in \linearExtensions(P)$.
In other words, any non-inversion of~$\omega$ is a non-inversion of~$\pi$, so that~${\pi \le \omega}$.
\end{proof}

@@

\begin{proof}
We just need to show that the poset~$\contactLess{P}$ satisfies the conditions of \cref{prop:WOIP}.
Consider~$i < j < k$ such that~$i \contactLess{P} k$.
If~$\omega^{-1}(i) < \omega^{-1}(j)$, then~$i \contactLess{P} j$ by \cref{lem:consequenceRectangle1}.
Similarly, if~$\omega^{-1}(j) < \omega^{-1}(k)$, then~$j \contactLess{P} k$ by \cref{lem:consequenceRectangle1}.
We can thus assume that~${\omega^{-1}(i) > \omega^{-1}(j) > \omega^{-1}(k)}$.
Decompose the triangular shape into three regions: the region~$A$ of all points located northeast of the last elbow of the pipe~$j$ of~$P$, the region~$B$ of all points located northwest or southeast of an elbow of the pipe~$j$ of~$P$, and the region~$C$ of all points located southwest of the first elbow of the pipe~$j$ of~$P$.
Since~$i \contactLess{P} k$, there is a path~$\pi$ form the exiting point of the pipe~$i$ of~$P$ to the entering point of the pipe~$j$ of~$P$ which travels along the pipes of~$P$, possibly jumping from the northwest pipe to the southeast pipe of a contact it encounters.
Since~$i < j$ and~$\omega^{-1}(j) > \omega^{-1}(k)$, the path~$\pi$ starts in region~$A$ and ends in region~$C$, so that it necessarily passes from region~$A$ to region~$C$.
Since the southwest corner of~$A$ is located northeast of the northeast corner of~$C$, this forces an elbow~$e$ of~$\pi$ to lie in region~$B$.
\cref{lem:rectangle} then ensures that either~$i \contactLess{P} j$ (if~$e$ is above the pipe~$j$), or~$j \contactLess{P} k$ (if~$e$ is below the pipe~$j$).
The proof is similar if~$i \contactMore{P} k$.
\end{proof}

@@

\begin{lemma}
\label{lem:flippable}
Consider two acyclic pipe dreams~$P,P' \in \acyclicPipeDreams(\omega)$ connected by the flip of a contact between their pipes~$i$ and~$j$.
Then any directed path in~$P\contact$ or~$P'{}\contact$ between~$i$ and~$j$ is an arc.
\end{lemma}

@@

\vincent{This section is in progress}
In this section, we describe three algorithms to decide whether two permutations~$\pi$ and~$\pi'$ are equivalent for the pipe dream congruence~$\equiv_\omega$.
The first two algorithms actually construct the unique pipe dream~$P \in \acyclicPipeDreams(\omega)$ of which~$\pi$ is a linear extension, while the last algorithm uses a rewriting technique.

@@

Our second algorithm is an \defn{insertion algorithm} that constructs the unique pipe dream~$P \in \acyclicPipeDreams(\omega)$ of which~$\pi$ is a linear extension.
It is inspired from~\cite{Pilaud-BrickAlgebra} and is similar to the insertion in binary search trees.

\begin{proposition}
\label{prop:insertionAlgorithm}
For any permutations~$\pi,\omega \in \fS_n$ such that~$\pi < \omega$ in weak order, the unique pipe dream~$P \in \acyclicPipeDreams(\omega)$ such that~$\pi \in \linearExtensions(P)$ can be constructed starting from the empty triangular shape and inserting the pipes~$\pi(1), \dots, \pi(n)$ one by one in the order of the permutation~$\pi$ as northwest as possible.
More precisely, at step~$t$, we consider the set~$X_t$ of free southeast elbows in the rectangle~$[\pi(t)] \times [\omega^{-1}(\pi(t))]$, and we insert a pipe starting at row~$\pi(t)$, ending at column~$\omega^{-1}(\pi(t))$, covering each elbow of~$X_t$, and adding a new free southeast elbow in between any two consecutive elbows of~$X_t$
See \cref{fig:insertionAlgorithm}.

\begin{figure}
	\centerline{\includegraphics[scale=.9]{insertion}}
	\caption{Insertion algorithm for the permutations~$\pi = 513264$ and~$\omega = 561324$.}
	\label{fig:insertionAlgorithm}
\end{figure}
\end{proposition}

\vincent{I am still working below this line.}
We need to argue that this algorithm indeed creates a pipe dream of~$\acyclicPipeDreams(\omega)$.
To see it, we observe that the following invariant is maintained throughout the algorithm.
